---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(tmap)
```

# cbssuitabilityhaiti

<!-- badges: start -->
DOI
R-CMD-Check
<!-- badges: end -->

# Overview

This packages combines data collected as part of ?????? conducted in Haiti. The projects were 
supported by SOIL at ??????.

## Installation

You can install the development version of cbssuitabilityhaiti from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/cbssuitabilityhaiti")
```

Alternatively, you can download the individual datasets as a CSV or XLSX file 
from the table below.

```{r, echo=FALSE}

extdata_path <- "https://github.com/openwashdata/cbssuitabilityhaiti/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  mutate(file_name = str_remove(file_name, ".rda")) |> 
  rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```


# Projects

## mWater

(One sentence) 
Evaluating the potential of ........

### Description

(One to two short paragraphs)
This Project focuses on determining .......

### Research Question

(One to two questions)

### Data 

Description of the Data. When/Where it was collected. For how long.

The package provides access one data set.

```{r}
library(cbssuitabilityhaiti)
```

```{epoxy}
The `mwater` data set has {ncol(mwater)} variables and {nrow(mwater)} observations. For an overview of the variable names, see the following table.  

```

```{r, eval=FALSE}
mwater
```


```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "mwater.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

A small visualisation of this data set

```{r, echo=FALSE}
qtm(mwater)
```


## Okap

(One sentence) 
Evaluating the potential of ........

### Description

(One to two short paragraphs)
This Project focuses on determining .......

### Research Question

(One to two questions)

### Data 

Description of the Data. When/Where it was collected. For how long.

The package provides access one data set.

```{r}
library(cbssuitabilityhaiti)
```

```{epoxy}
The `okap` data set has {ncol(okap)} variables and {nrow(okap)} observations. For an overview of the variable names, see the following table.  

```

```{r, eval=FALSE}
okap
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::filter(file_name == "mwater.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

A small visualisation of this data set

```{r, echo=FALSE}
qtm(okap)
```


## Examples

This is a basic example which shows you how to use the data:

```{r example}
library(cbssuitabilityhaiti)
## basic example code
```
